\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-04-23 intro.tex
updates to reviews

In this chapter, we propose an alternative view to decision-making
using in-situ automated Planning \cite{ghallab04}. More importantly,
we demonstrate that earlier control paradigms (primarily driven by
\emph{reactive} approaches to control) can be augmented to provide
targeted observation and sampling as well as a more nuanced and
balanced consideration of mission objectives, environmental conditions
and available resources. Planning (or deliberation or projection in
action space) is important to balance current needs of a robot with
future desires or goals taking into account time and available
resources in-situ on the robot. Planning without time or action
planning in and of itself is not adequate; dealing with time ensures
that the balance between now and the future is handled
systematically. Yet, plans have to be tied to robotic action and in
the real-world which is not static. Further, plans project future
state which is not only likely to change, but is based on substantial
uncertainty\footnote{It is important to remember that these principles
  are broad much as Dwight Eisenhower is reputed to have said
  ``\emph{In preparing for battle I have always found that plans are
    useless, but planning is indispensable}'' and ``\emph{Failing to
    plan is planning to fail}''}. Use of environmental cues is often
hampered by poor predicitve skill of ocean models and little to no
availability of a priori synoptic views of the survey area

@@

Criticism of AI planning, on the other hand, also well articulated by
Brooks, was targeting the performance of in-situ
deliberation. Terrestrial robotic platforms were well-known to
frequently stop and ``plan''; assuming in the meantime that the world
will remain static. In the marine robotics world, such a situation
would simply not be permissible given the need to make continuous
observations without stopping an under-articulated robotic platform
such as an AUV. However, algorithmic and implementational advances by
the AI community over the years have produced a range of planners with
demonstrated embedded capabilities
\cite{simmons94,Haigh98,alami:1998p820,chien00,mus98,teichteil07}. Our
work has been influenced by and in turn influenced a wide range of
activities in the field of AI Planning and Plan Execution with
demonstrated real-world and highly visible capabilities in the Space
domain \cite{mus98,rajan00,aichang04,bresina05} for NASA. Lessons
learned have been applied to the world of marine robotics which we
bring together in this chapter. Important lessons that were derived
from this rich legacy included the following:

@@

We have developed, tested and deployed the Teleo-Reactive EXecutive
(\rx) an on-board adaptive control system that integrates AI based
planning and state estimation in a hybrid
executive\cite{mcgann08a,mcgann08b,py10}.  State estimation allows the
system to keep track of world evolution as time advances based on a
formal model distributed through the architecture.  Onboard planning
and execution enables adaptation of navigation and instrument control
based on the state identified during state estimation.  It further
enables goal-directed commanding within the context of projected
mission state and allows for replanning for off-nominal situations and
opportunistic science events. The framework in addition to being used
on an AUV, is general enough be used for controlling a personal robot
\cite{pr2,Meeussen:2010dn,mcgann2009} and is deployed on a European
planetary rover testbed \cite{goac11}. While it is possible to
integrate probabilistic estimation on top of this framework
\cite{mcgann08d}, such representation is not explicitly handled by the
framework and is out of scope of this chapter.  Deliberation and
reaction are integrated systematically over different temporal and
functional scopes within a single agent and a single model that covers
the needs of high-level mission management, low-level navigation,
instrument control, and detection of unstructured and poorly
understood phenomena. \rx is deployed on MBARI's \emph{Dorado} AUV
shown in Fig. \ref{fig:auv-fig}, which to the best of our knowledge is
the only operational marine platform anywhere being used for routine
scientific surveys with onboard plan synthesis.

>>>>

In this chapter, we propose an alternative view to decision-making
using in-situ automated Planning \cite{ghallab04}. More importantly,
we demonstrate that earlier control paradigms (primarily driven by
\emph{reactive} approaches to control) can be augmented to provide
targeted observation and sampling as well as a more nuanced and
balanced consideration of mission objectives, environmental conditions
and available resources. Planning (or deliberation or projection in
action space \kcomment{and used interchangeably in this chapter}) is
important to balance current needs of a robot with future desires or
goals taking into account time and available resources in-situ on the
robot. Planning without time or action planning in and of itself is
not adequate; dealing with time ensures that the balance between now
and the future is handled systematically. Yet, plans have to be tied
to robotic action and in the real-world which is not static. Further,
plans project future state which is not only likely to change, but is
based on substantial uncertainty\footnote{It is important to remember
  that these principles are broad much as Dwight Eisenhower is reputed
  to have said ``\emph{In preparing for battle I have always found
    that plans are useless, but planning is indispensable}'' and
  ``\emph{Failing to plan is planning to fail}''}. Use of
environmental cues is often hampered by poor \kcomment{predictive}
skill \cite{anderson2009} of ocean models and little to no
availability of a priori synoptic views of the survey
area. \kcomment{Planning therefore provides a ``buffer'' against the
  stochastic environment.}

@@

Criticism of AI planning, on the other hand, also well articulated by
Brooks, was targeting the performance of in-situ
deliberation. Terrestrial robotic platforms were well-known to
frequently stop and ``plan''; assuming in the meantime that the world
will remain static. In the marine robotics world, such a situation
would simply not be permissible given the need to make continuous
observations without stopping a \kcomment{non-holonomic} robotic
platform such as an AUV. However, algorithmic and implementational
advances by the AI community over the years have produced a range of
planners with demonstrated embedded capabilities
\cite{simmons94,Haigh98,alami:1998p820,chien00,mus98,teichteil07}. Our
work has been influenced by and in turn influenced a wide range of
activities in the field of AI Planning and Plan Execution with
demonstrated real-world and highly visible capabilities in the Space
domain \cite{mus98,rajan00,aichang04,bresina05} for NASA. Lessons
learned have been applied to the world of marine robotics which we
bring together in this chapter. Important lessons that were derived
from this rich legacy included the following:

@@

We have developed, tested and deployed the Teleo-Reactive EXecutive
(\rxe) an on-board adaptive control system that integrates AI based
planning and state estimation in a hybrid
executive\cite{mcgann08a,mcgann08b,py10}.  State estimation allows the
system to keep track of world evolution as time advances based on a
formal model distributed through the architecture.  Onboard planning
and execution enables adaptation of navigation and instrument control
based on the state identified during state estimation.  It further
enables goal-directed commanding within the context of projected
mission state and allows for replanning for off-nominal situations and
opportunistic science events. The framework in addition to being used
on an AUV, is general enough \kcomment{to} be used for controlling a
personal robot \cite{pr2,Meeussen:2010dn,mcgann2009} and is deployed
on a European planetary rover testbed \cite{goac11}. While it is
possible to integrate probabilistic estimation on top of this
framework \cite{mcgann08d}, such representation is not explicitly
handled by the framework and is out of scope of this chapter.
Deliberation and reaction are integrated systematically over different
temporal and functional scopes within a single agent and a single
model that covers the needs of high-level mission management,
low-level navigation, instrument control, and detection of
unstructured and poorly understood phenomena. \rx is deployed on
MBARI's \emph{Dorado} AUV shown in Fig. \ref{fig:auv-fig}, which to
the best of our knowledge is the only operational marine platform
anywhere being used for routine scientific surveys with onboard plan
synthesis.

### 2012-04-23 results.tex
updates to reviews

A significant driver of the applicability of our work was the growing
need for AUV missions to be driven by sensor data. This trend is
illustrated by the implementation of MBARI's \emph{Gulper}
water-samplers \cite{Bird07}, $2$ litre syringe style containers $10$
of which are encased within the mid-body section of our Dorado vehicle
as shown in Fig \ref{fig:gulper}. However, the original design and
intended use of operation continued to be dominated by a naive belief
of collecting and returning water samples as traditional ship-borne
methods often employ, namely using pre-scripted methods. With the
installation of \rx onboard the Dorado with its capability sense the
environment and to appropriately synthesize plans and replan on the
fly, more imaginative methods of sample acquisition were targeted.

>>>>

A significant driver of the applicability of our work was the growing
need for AUV missions to be driven by sensor data. This trend is
illustrated by the implementation of MBARI's \emph{Gulper}
water-samplers \cite{Bird07}, $2$ litre syringe style containers $10$
of which are encased within the mid-body section of our Dorado vehicle
as shown in Fig \ref{fig:gulper}. However, the original design and
intended use of operation continued to be dominated by a naive belief
of collecting and returning water samples as traditional ship-borne
methods often employ, namely using pre-scripted methods. With the
installation of \rx onboard the Dorado with its capability sense the
environment and to appropriately synthesize plans and replan on the
fly, more imaginative methods of sample acquisition were targeted.
\kcomment{The Dorado uses two onboard computers: a main vehicle
  computer which is a 244 Mhz PC/104 stack running the QNX real-time
  operating system, and a 367 MHz EPIC EPX-GX500 AMD Geode stack
  running Linux and \rxe. A socket-based communication protocol allows
  the exchange of goals and state updates between \rx and the lower
  level functional layer.}\comment{We should say something about the
  peak load during re-planning.}
